Beef Pricing Dynamics in Canadian Supermarkets

### 2024-12-03 paper/paper.qmd
Add files via upload

@@

## Overview
The data for this study came from Project Hammer[@PG].The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help of `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed] and `kableExtra` [@citeKableExtra].The raw data covers eight major large supermarkets in Canada and the changes in their products between February 28, 2024, and November 26, 2024.

@@

@@

The outcome variable, current price, represents the most recent market price of products in the dataset. It serves as a critical metric for understanding pricing strategies, consumer behavior, and market trends. By analyzing current_price, we can evaluate the impact of factors such as discounts (old price), vendor differences, and temporal trends (e.g., month) on product pricing.The line graph [@fig-current] illustrates the average monthly current_price trend across the dataset. The most notable pattern is the significant drop in prices between months 6 and 8, reaching the lowest average in month 8. This is followed by a steady upward trend, with prices gradually increasing and peaking in month 11.

This trend likely reflects seasonal or market dynamics that influence product pricing. The decline in prices during the mid-year months could indicate promotional events, seasonal sales, or a decrease in consumer demand during that period. Conversely, the rise in prices from months 9 to 11 may correspond to increased demand, possibly due to holiday preparations or end-of-year market adjustments.

@@

## Predictor variables

@@

**Old price**:This variable represents the product's previous price. It serves as an indicator of discounts or price changes, which are key factors in determining the current_price. It provides a direct measure of pricing adjustments over time.

@@

I also processed the dates, focusing primarily on monthly changes. I simplified the nowtime variable to retain only the month. Additionally, some data in the dataset contained missing values. However, when reviewing the raw data, I observed that for the same product within the same month, the current_price and old_price were relatively consistent. For example, if there was complete data on the 1st of the month, missing data on the 4th, and complete data again on the 10th, the values on the 1st and the 10th were identical. Therefore, to control the sample size, I chose to directly remove entries with missing values.After handling this, it will not have a significant impact on my monthly figures.

@@

## Collection methods

@@

In this setup, the primary expense will be software development to automate the data collection and storage process. This system will help streamline data gathering and ensure consistency over time. However, all activities must be conducted with the explicit consent of the supermarkets and in full compliance with ethical and legal standards to ensure no infringement on the rights of individuals or groups.This way, we can obtain three pieces of data: the product name, where it is sold, and its price.We can also collaborate with these supermarkets to obtain their beef product pricing information directly.

>>>>

## Estimand

We want to estimate our current price using vendor, month, and old price.After a quick review of the raw data, I believe that the estimated current price should ultimately be lower than the old price.

@@

## Overview
The data for this study came from Project Hammer[@PG].The datasets were cleaned and analysed using the statistical programming software `R` [@citeR] along with the help of `tidyverse` [@citeTidyverse], `knitr` [@citeKnitr], `ggplot2` [@citeGgplot], `here` [@citeHere], `dplyr` [@citeDplyr], `rstanarm` [@citeRstanarm], `arrow` [@citeArrow], `broom.mixed` [@citeBroomMixed] and `kableExtra` [@citeKableExtra].The raw data covers eight major large supermarkets in Canada and the changes in their products between February 28, 2024, and November 26, 2024.In the following section, I will explain the components of my dataset, the time period it covers, and other detailed information.

@@

- **Nowtime:** Timestamp indicating when the data was gathered.

- **Vendor:** 	One of the 7 grocery vendors.

- **Product Id:** An internal ID for a product - it represents 1 unique product at 1 vendor. It will allow you to track prices on the same product over time.

- **Product Name:** Product name. May include Brand, may include Units in it.

- **Brand:** Brand name. May be blank for some vendors.

- **Units:** Units (grams, kg, number of items in package). May be blank for some vendors/products.

- **Current Price:** 	Price at time of extract.

- **Old Price:** An "old" struck-out price. This indicates that there was a Sale on. This is how you can differentiate a price-drop advertised as a Sale vs. a "quiet" decrease without a Sale.

- **Price Per Unit:** Price per unit, as shown on the vendor website. May not be consistent with real calculation of "current price" divided by "units".

- **Other:** 	Other details that appeared on the listing. Values may be "Out of stock", "SALE", "Best seller", "$5.00 MIN 2".

@@

The outcome variable, current price, represents the most recent market price of products in the dataset. It serves as a critical metric for understanding pricing strategies, consumer behavior, and market trends. By analyzing current price, we can evaluate the impact of factors such as discounts (old price), vendor differences, and temporal trends (e.g., month) on product pricing.The line graph [@fig-current] illustrates the average monthly current price trend across the dataset. The most notable pattern is the significant drop in prices between months 6 and 8, reaching the lowest average in month 8. This is followed by a steady upward trend, with prices gradually increasing and peaking in month 11.

This trend likely reflects seasonal or market dynamics that influence product pricing. The decline in prices during the mid-year months could indicate promotional events, seasonal sales, or a decrease in consumer demand during that period. Conversely, the rise in prices from months 9 to 11 may correspond to increased demand, possibly due to holiday preparations or end of year market adjustments.

@@

The summary shows that the current_price in the dataset ranges from 0.39 to 101.15, with most prices concentrated in the lower range, as indicated by a median of 6.87 and a third quartile of 10.99. The average price is 9.56, while the standard deviation of 11.39 suggests significant variability, likely influenced by a few high outliers.
```{r}
#| label: tbl-summary
#| fig-cap: Summary Statistics for Current Price
#| echo: false

summary_stats <- summary(cleaned_data$current_price)
std_dev <- sd(cleaned_data$current_price, na.rm = TRUE)

stats_table <- data.frame(
  Statistic = c("Minimum", "1st Quartile", "Median", "Mean", "3rd Quartile", "Maximum", "Standard Deviation"),
  Value = c(
    summary_stats["Min."],
    summary_stats["1st Qu."],
    summary_stats["Median"],
    summary_stats["Mean"],
    summary_stats["3rd Qu."],
    summary_stats["Max."],
    std_dev
  )
)

kable(stats_table)

```
## Predictor variables

@@

**Old price**:This variable represents the product's previous price. It serves as an indicator of discounts or price changes, which are key factors in determining the current price. It provides a direct measure of pricing adjustments over time.

@@

I also processed the dates, focusing primarily on monthly changes. I simplified the nowtime variable to retain only the month. Additionally, some data in the dataset contained missing values. However, when reviewing the raw data, I observed that for the same product within the same month, the current price and old price were relatively consistent. For example, if there was complete data on the 1st of the month, missing data on the 4th, and complete data again on the 10th, the values on the 1st and the 10th were identical. Therefore, to control the sample size, I chose to directly remove entries with missing values.After handling this, it will not have a significant impact on my monthly figures.

@@

## idealize Collection methods

@@

The primary expense in this setup will be software development to automate the data collection and storage process. This system will streamline data gathering and ensure consistency over time. To achieve this, all activities must be conducted with the explicit consent of the supermarkets and in full compliance with ethical and legal standards, ensuring no infringement on the rights of individuals or groups. Through this system, we can obtain three key pieces of data: the product name, its location of sale, and its price. Additionally, we can collaborate with supermarkets to directly access their beef product pricing information. By complementing this with daily sales data, we can better analyze the demand for beef products.Building on this, we will divide the analysis into two major sections: rural and urban areas. The datasets for these two categories will be analyzed separately, as the demand and market size for beef products differ significantly between rural and urban areas. This approach will allow us to conduct a more targeted analysis of each group.

## Survey Questionnaire

With the market survey and data collection complete, we will have gathered enough information to proceed. Next, our focus will shift to understanding the people themselves. By analyzing demand for beef products across different demographics—such as age, region, and gender—we aim to study how demand varies among different groups. This will allow us to gain deeper insights into consumer preferences and behaviors.

Section 1: Personal Information

1. Gender

 - Male
 - Female
 - Other/Prefer not to say

2. Age Group

 - Under 18
 - 18-24
 - 25-34
 - 35-44
 - 45-54
 - 55-64
 - 65 and above

3. Location

- Province/Territory:
(Drop-down menu or text input for respondents to specify their province/territory in Canada)

4. Area:
 - Urban
 - Rural

Section 2: Beef Product Demand

1.How would you describe your demand for beef products?

 - High demand (consume beef frequently, buy large quantities)
 - Moderate demand (consume beef occasionally, buy moderate quantities)
 - Low demand (rarely consume beef, buy small quantities)
 - No demand (do not consume beef)

2.How often do you purchase beef products?

 - Daily
 - 2-3 times a week
 - Weekly
 - Biweekly
 - Monthly
 - Rarely/Never

Section（Optional） 3： Additional Feedback

1.Are there any factors that influence your beef product purchasing decisions?

 - (e.g., price, availability, dietary preferences, cultural practices, etc.)

2.Would you like to be contacted for future surveys or updates on this research?

 - Yes
 - No
